Deliminative verbal reduplication in Mandarin Chinese

### 2025-06-16 chapters/ch01-Intro.tex
check main text before sec 2.3

The data in this paper was drawn from several sources.
In addition to introspection, the Modern Chinese subcorpus of the corpus of the \textit{Center for Chinese Linguistics of Peking University} ({CCL}) \citep{Zhanetal2003, Zhanetal2019} and the BCC corpus \citep{BCC} was also consulted. 
Further, examples from novels and plays written by native speakers were considered.
Corpus data provides natural and contextualized examples, and contains a variation of linguistic properties \citep[921]{MM2009a}.
This can help us discover relevant constraints that can otherwise go unnoticed through introspection.

>>>>

The data in this paper was drawn from several sources.
In addition to introspection, the Modern Chinese subcorpus of the corpus of the \textit{Center for Chinese Linguistics of Peking University} ({CCL}; \citealt{Zhanetal2003, Zhanetal2019}) and the BCC corpus \citep{BCC} were also consulted. 
Further, examples from novels and plays written by native speakers were considered.
Corpus data provides natural and contextualized examples, and contains a variation of linguistic properties \citep[921]{MM2009a}.
This can help us discover relevant constraints that can otherwise go unnoticed through introspection.

### 2025-06-16 chapters/ch02-Phenomenon.tex
check main text before sec 2.3

		\ex \gll \objex{shuo1-kan4-kan4}\\
		 say-look-look\\ \jambox{A-\obj{kan4}-\obj{kan4}\footnotemark}
         \glt `try to say a little bit'
		\z
\footnotetext{One reviewer points out that this form is not acceptable for him/her.
The acceptability of the A-\obj{kan4-kan4} form seems to vary among Mandarin Chinese speakers.
\citet[73]{Cheng2012} suggests that unlike AA-\obj{kan4}, A-\obj{kan4-kan4} is an emerging construction that is still undergoing grammaticalization.
He found in the Academia Sinica Balanced Corpus of Modern Chinese (Sinica Corpus; \citealt{sinica}) 23 tokens of A-\obj{kan4-kan4} but 141 tokens of AA-\obj{kan4} (p. 64).
This shows that the former is less common than the latter.
(\ref{ex:Akankan}) is an example of A-\obj{kan4-kan4} from the Sinica Corpus.
\ea\label{ex:Akankan} %就順著北風的吹勢，試看看能否將海沙攔下來吧！
\gll \objex{jiu4} \objex{shun4-zhe} \objex{bei3-feng1-de} \objex{chui1-shi4}, \obj{shi4-kan4-kan4} \objex{neng2} \objex{fou3} \objex{jiang1} \objex{hai3-sha1} \objex{lan2-xia4lai2} \objex{ba}!\\
just follow-\textsc{prog} north-wind-\textsc{de} blow-direction try-look-look can not make sea-sand block-down \textsc{ptc}\\\jambox{(Sinica Corpus)}
\glt `Let's just follow the north wind, try and see if we can block the sand!'
\z

@@

\ex\label{ex:forms-di} for disyllabic verbs: \obj{lai2-wang3} come-go `come and go/communicate'
		\ea \gll \objex{lai2-wang3-lai2-wang3}\\
		come-go-come-go\\ \jambox{ABAB}
        \glt `communicate a little bit'

		\ex \gll \objex{lai2-wang3-le-lai2-wang3}\\
		come-go-\textsc{pfv}-come-go\\ \jambox{AB-\textit{le}-AB}
        \glt `communicated a little bit'

		\ex \gll \objex{lai2-lai2-wang3-wang3}\\
		come-come-go-go\\ \jambox{AABB}
        \glt `coming and going'
		\z

@@

 \citet{Fan1964}, \citet{Arcodiaetal2014}  and \citet{Xie2020} compare the AA, ABAB and AABB forms of reduplication 
and found a number of differences between the AA, ABAB forms compared to the AABB form in terms of their semantics, productivity, syntactic distribution and origin. 
Specifically, \citet[17--18]{Arcodiaetal2014}, \citet[144]{MelloniBasciano2018} and \citet[90]{Xie2020} identify that AA and ABAB have a diminishing meaning, 
namely that the event happens for a short duration or to a small extent. 
By contrast, AABB expresses an increasing meaning, which indicates a repetition or an action in progress. 
\citet[Sec.\,3.1]{Xie2020} also finds that AA and ABAB have relatively high productivity, 
whereas the productivity of AABB is low. 
She further shows that AABB is generally correlated with the lack of a postverbal object, but the direct object remains present when a transitive verb undergoes AA or ABAB patterns of reduplication. 
\citet[277]{Fan1964} proposes that AA, ABAB originated from the verb-measure word combination from Middle Chinese, 
while AABB developed from the reiterative rhetoric from Old Chinese. 
These differences seem to suggest that there is a fundmental difference between these two groups of verbal reduplication. 
The current study will only focus on the AA, A-\textit{yi}-A, A-\textit{le}-A, A-\textit{le}-\textit{yi}-A, ABAB and AB-\textit{le}-AB forms of verbal reduplication in Mandarin Chinese.
AA-\obj{kan4}, A-\obj{kan4}-\obj{kan4}, AAB, A-\textit{yi}-AB, A-\textit{le}-AB will also be mentioned occasionally to provide further arguments.
In what follows, the term \textit{reduplication} will be used  to refer specifically to the AA, A\hyp{}\textit{yi}\hyp{}A, A-\textit{le}-A, A-\textit{le}-\textit{yi}-A, ABAB and AB-\textit{le}-AB forms, if not specified otherwise.

@@

\ex[*]{\gll \objex{zhe4} \objex{ge4} \objex{wen4ti2} \objex{da4jia1} \objex{hai2} \objex{yao4} \objex{tao3lun4-tao3lun4} \objex{yi2} \objex{hui4er}/\objex{ji3} \objex{ci4}.\\
    this \textsc{clf} question everybody still need.to discuss-discuss one while/several time\\}
%    \glt `As for this question, everybody still needs to discuss it for a while/several times.'}
\z\z
As we will show in \sectref{sec:adjuncts}, this can be explained by the redundant semantics of the reduplication and duration and frequency phrases.
%with an expression that quantifies the duration or the extent of the event described by the sentence due to its semantic properties (see \sectref{sec:adjuncts} for a detailed discussion).

Thirdly, \citet[319]{SuiHu2016} note that reduplicated verbs cannot be aspect marked (\ref{ex:redup-asp}).
\ea\label{ex:redup-asp}
\ea[]{ \gll \objex{wo3} \objex{kan4-le/guo4/zhe} \objex{zhe4} \objex{ben3} \objex{shu1}\\
I read-\textsc{pfv/exp/dur} this \textsc{clf} book\\
\glt `I read/have read/am reading this book.'
}
\ex[*]{ \gll \objex{wo3} \objex{kan4-kan4-le/guo4/zhe} \objex{zhe4} \objex{ben3} \objex{shu1}\\
    I read-read-\textsc{pfv/exp/dur} this \textsc{clf} book\\
%    \glt `I read/have read/am reading this book.'
}
\z\z
However, as shown in (\ref{ex:xiaole}), \obj{le} can occur in between the reduplicated verb
 and expresses the perfective aspectual meaning, that the event denoted by the sentence is realized.
Thus, we argue that the reduplication can in fact be aspect marked,
but only by \obj{le} `\textsc{pfv}' but not the other aspect markers.
As we will argue in \sectref{sec:aspM}, this is also not a syntactic but a semantic constraint.
%, though, except with the perfective aspect marker \textit{le} (see \sectref{sec:aspM}).% for further discussions on interactions with aspect markers).

@@

\ex\label{ex:neg2}
% 所谓职业的读书者，譬如学生因为升学，教员因为要讲功课，不翻翻书，就有些危险的就是。(CCL 1920s\1928\文学\1928 鲁迅 而已集.txt)
\gll \objex{jiao4yuan2} \ldots\, \objex{bu4} \obj{fan1-fan1} \objex{shu1}, \objex{jiu4} \objex{you3xie1} \objex{wei1xian3} \objex{de} \objex{jiu4} \objex{shi4}.\\
lecturer {} not flip-flip book just somewhat dangerous \textsc{de} just be\\ \jambox{(CCL)}
\glt `If lecturers do not read books, then it is a bit dangerous.'

@@

\ex\label{ex:qobj2}% 男人跟男人的沟通比较快，打打球、喝喝两 杯就行了。（BCC 微博）
\gll \objex{nan2ren2} \objex{gen1} \objex{nan2ren2} \objex{de} \objex{gou1tong1} \objex{bi3jiao4} \objex{kuai4}, \ldots\, \obj{he1-he1} \objex{liang3} \objex{bei1} \objex{jiu4} \objex{xing2} \objex{le}.\\
man and man \textsc{de} communication relatively fast {} drink-drink two cup just ok \textsc{ptc}\\ \jambox{(BCC)}
\glt `The communication between men is relatively fast, \ldots\, they just need to drink two cups together.'

@@

\noindent
The ungrammaticality of (\ref{ex:vp}) can be explained by the fact that almost all Mandarin Chinese adverbials are obligatorily pre-verbal \citep[50]{Ernst2014}, as illustrated in the example below in (\ref{ex:zouadv}).
\ea\label{ex:zouadv}
\ea[]{\gll \objex{ta1} \objex{qiao1qiao1de} \objex{zou3-le}\\
    he quietly go-\textsc{pfv}.\\
    \glt `He quietly went away.'
}
\ex[*]{\gll \objex{ta1} \objex{zou3-le} \objex{qiao1qiao1de}.\\
    he go-\textsc{pfv} quietly\\
}
\z\z

\citet[50]{Ernst2014} mentions that the only possible post-verbal adverbials are participant PPs (\textit{with somebody}),
manner or resultative V-de constructions as well as 
duration and frequency expressions.
(\ref{ex:inVCPa}) shows a case of inserting a frequency expression into a VP.
This suggests that it is possible to insert adverbials into a VP.

@@

\ex  \gll \objex{kai1}  \objex{san1} \objex{ci4} \objex{men2}\\
open three time door\\
\glt `open the door three times'
\z\z
For reduplication, this is also not possible (\ref{ex:inVV}), as the object cannot be inserted between the two verbs.
\ea\label{ex:inVV}
\ea[]{\gll \obj{kai1}-(\obj{le})-\obj{kai1} \objex{men2}\\
    open-\textsc{pfv}-open door\\
    \glt `open the door for a little while'}

>>>>

		\ex \gll \objex{shuo1-kan4-kan4}\\
		 say-look-look\\ \jambox{A-\obj{kan4}-\obj{kan4}\footnotemark}
         \glt `try to say a little bit'
		\z
\footnotetext{One reviewer points out that this form is not acceptable for him/her.
The acceptability of the A-\obj{kan4-kan4} form seems to vary among Mandarin Chinese speakers.
\citet[73]{Cheng2012} suggests that unlike AA-\obj{kan4}, A-\obj{kan4-kan4} is an emerging construction that is still undergoing grammaticalization.
He found in the Academia Sinica Balanced Corpus of Modern Chinese (Sinica Corpus; \citealt{sinica}) 23 tokens of A-\obj{kan4-kan4} but 141 tokens of AA-\obj{kan4} (p. 64).
This shows that the former is less common than the latter.
(\ref{ex:Akankan}) below is an example of A-\obj{kan4-kan4} from the Sinica Corpus.
\ea\label{ex:Akankan} %就順著北風的吹勢，試看看能否將海沙攔下來吧！
\gll \objex{jiu4} \objex{shun4-zhe} \objex{bei3-feng1-de} \objex{chui1-shi4}, \obj{shi4-kan4-kan4} \objex{neng2} \objex{fou3} \objex{jiang1} \objex{hai3-sha1} \objex{lan2-xia4lai2} \objex{ba}!\\
just follow-\textsc{dur} north-wind-\textsc{de} blow-direction try-look-look can not make sea-sand block-down \textsc{ptc}\\\jambox{(Sinica Corpus)}
\glt `Let's just follow the north wind, try and see if we can block the sand!'
\z

@@

\ex\label{ex:forms-di} for disyllabic verbs: \obj{lai2-wang3} come-go `come and go/communicate'
		\ea\label{ex:forms-dia} \gll \objex{lai2-wang3-lai2-wang3}\\
		come-go-come-go\\ \jambox{ABAB}
        \glt `communicate a little bit'

		\ex\label{ex:forms-dib} \gll \objex{lai2-wang3-le-lai2-wang3}\\
		come-go-\textsc{pfv}-come-go\\ \jambox{AB-\textit{le}-AB}
        \glt `communicated a little bit'

		\ex\label{ex:forms-dic} \gll \objex{lai2-lai2-wang3-wang3}\\
		come-come-go-go\\ \jambox{AABB}
        \glt `coming and going'
		\z

@@

 \citet{Fan1964}, \citet{Arcodiaetal2014}  and \citet{Xie2020} compare the AA, ABAB and AABB forms of reduplication 
and found a number of differences between the AA, ABAB forms compared to the AABB form in terms of their semantics, productivity, syntactic distribution and origin. 
Specifically, \citet[17--18]{Arcodiaetal2014}, \citet[144]{MelloniBasciano2018} and \citet[90]{Xie2020} identify that AA and ABAB have a diminishing meaning, 
namely that the event happens for a short duration or to a small extent. 
By contrast, AABB expresses an increasing meaning, which indicates a repetition or an action in progress
(compare (\ref{ex:forms-dia}) and (\ref{ex:forms-dic})). 
\citet[Sec.\,3.1]{Xie2020} also finds that AA and ABAB have relatively high productivity, 
whereas the productivity of AABB is low. 
She further shows that AABB is generally correlated with the lack of a postverbal object, but the direct object remains present when a transitive verb undergoes AA or ABAB patterns of reduplication. 
\citet[277]{Fan1964} proposes that AA, ABAB originated from the verb-measure word combination from Middle Chinese, 
while AABB developed from the reiterative rhetoric from Old Chinese. 
These differences seem to suggest that there is a fundmental difference between these two groups of verbal reduplication. 
The current study will only focus on the AA, A-\textit{yi}-A, A-\textit{le}-A, A-\textit{le}-\textit{yi}-A, ABAB and AB-\textit{le}-AB forms of verbal reduplication in Mandarin Chinese.
AA-\obj{kan4}, A-\obj{kan4}-\obj{kan4}, AAB, A-\textit{yi}-AB, A-\textit{le}-AB will also be mentioned occasionally to provide further arguments.
In what follows, the term \textit{reduplication} will be used  to refer specifically to the AA, A\hyp{}\textit{yi}\hyp{}A, A-\textit{le}-A, A-\textit{le}-\textit{yi}-A, ABAB and AB-\textit{le}-AB forms, if not specified otherwise.

@@

\ex[*]{\gll \objex{zhe4} \objex{ge4} \objex{wen4ti2} \objex{da4jia1} \objex{hai2} \objex{yao4} \obj{tao3lun4-tao3lun4} \objex{yi2} \objex{hui4er}/\objex{ji3} \objex{ci4}.\\
    this \textsc{clf} question everybody still need.to discuss-discuss one while/several time\\}
%    \glt `As for this question, everybody still needs to discuss it for a while/several times.'}
\z\z
As we will show in \sectref{sec:adjuncts}, this can be explained by the redundant semantics of the reduplication and duration and frequency phrases.
%with an expression that quantifies the duration or the extent of the event described by the sentence due to its semantic properties (see \sectref{sec:adjuncts} for a detailed discussion).

Thirdly, \citet[319]{SuiHu2016} note that reduplicated verbs cannot be aspect marked (\ref{ex:redup-asp}).
\ea\label{ex:redup-asp}
\ea[]{ \gll \objex{wo3} \objex{kan4-le/guo/zhe} \objex{zhe4} \objex{ben3} \objex{shu1}\\
I read-\textsc{pfv/exp/dur} this \textsc{clf} book\\
\glt `I read/have read/am reading this book.'
}
\ex[*]{ \gll \objex{wo3} \obj{kan4-kan4-le/guo/zhe} \objex{zhe4} \objex{ben3} \objex{shu1}\\
    I read-read-\textsc{pfv/exp/dur} this \textsc{clf} book\\
%    \glt `I read/have read/am reading this book.'
}
\z\z
However, as shown in (\ref{ex:xiaole}), \obj{le} can occur in between the reduplicated verb
 and expresses the perfective aspectual meaning, that the event denoted by the sentence is realized.
Thus, we argue that the reduplication can in fact be aspect marked,
but only by \obj{le} `\textsc{pfv}' but not the other aspect markers.
As we will argue in \sectref{sec:aspM}, this is also not a syntactic but a semantic constraint.
%, though, except with the perfective aspect marker \textit{le} (see \sectref{sec:aspM}).% for further discussions on interactions with aspect markers).

@@

\ex\label{ex:neg2}
% 所谓职业的读书者，譬如学生因为升学，教员因为要讲功课，不翻翻书，就有些危险的就是。(CCL 1920s\1928\文学\1928 鲁迅 而已集.txt)
\gll \objex{jiao4yuan2} \ldots\, \objex{bu4} \obj{fan1-fan} \objex{shu1}, \objex{jiu4} \objex{you3xie1} \objex{wei1xian3} \objex{de} \objex{jiu4} \objex{shi4}.\\
lecturer {} not flip-flip book just somewhat dangerous \textsc{de} just be\\ \jambox{(CCL)}
\glt `If lecturers do not read books, then it is a bit dangerous.'

@@

\ex\label{ex:qobj2}% 男人跟男人的沟通比较快，打打球、喝喝两 杯就行了。（BCC 微博）
\gll \objex{nan2ren2} \objex{gen1} \objex{nan2ren2} \objex{de} \objex{gou1tong1} \objex{bi3jiao4} \objex{kuai4}, \ldots\, \obj{he1-he} \objex{liang3} \objex{bei1} \objex{jiu4} \objex{xing2} \objex{le}.\\
man and man \textsc{de} communication relatively fast {} drink-drink two cup just ok \textsc{ptc}\\ \jambox{(BCC)}
\glt `The communication between men is relatively fast, \ldots\, they just need to drink two cups together.'

@@

\noindent
The ungrammaticality of (\ref{ex:vp}) can be explained by the fact that almost all Mandarin Chinese adverbials are obligatorily pre-verbal \citep[50]{Ernst2014}, as illustrated in the example below in (\ref{ex:zouadv}).
\ea\label{ex:zouadv}
\ea[]{\gll \objex{ta1} \objex{qiao1qiao1de} \objex{zou3-le}.\\
    he quietly go-\textsc{pfv}\\
    \glt `He quietly went away.'
}
\ex[*]{\gll \objex{ta1} \objex{zou3-le} \objex{qiao1qiao1de}.\\
    he go-\textsc{pfv} quietly\\
}
\z\z

\citet[50]{Ernst2014} mentions that the only possible post-verbal adverbials are participant PPs (\textit{with somebody}),
manner or resultative V-de constructions as well as 
duration and frequency expressions.
(\ref{ex:inVCPb}) shows a case of inserting a frequency expression into a VP.
This suggests that it is possible to insert adverbials into a VP.

@@

\ex\label{ex:inVCPb}  \gll \objex{kai1}  \objex{san1} \objex{ci4} \objex{men2}\\
open three time door\\
\glt `open the door three times'
\z\z
For reduplication, this is also not possible (\ref{ex:inVV}), as the object cannot be inserted between the two verbs.
\ea\label{ex:inVV}
\ea[]{\gll \obj{kai1}-(\obj{le})-\obj{kai1} \objex{men2}\\
    open-\textsc{pfv}-open door\\
    \glt `open the door for a little while'}
